Explicit B-spline regularization in diffeomorphic image registration

### 2013-10-25 paper.tex
ENH:  More edits.

antsRegistration --dimensionality 3 \
                 --output ${prefix} \
                 --use-histogram-matching 1 \
                 --initial-moving-transform [${fixed},${moving},1] \
                 --transform Rigid[0.1] \
                 --metric MI[${fixed},${moving},1,32,Regular,0.25] \ 
                 --convergence 1000x500x250 \
                 --smoothing-sigmas 2.0x1.0x0.0mm \
                 --shrink-factors 4x3x2 \
                 --transform Affine[0.1] \
                 --metric MI[${fixed},${moving},1,32,Regular,0.25] \ 
                 --convergence 1000x500x250 \
                 --smoothing-sigmas 2.0x1.0x0.0mm \
                 --shrink-factors 4x3x2 \
                 --transform BSplineSyN[0.1,10x10x10,0x0x0,3] \
                 --metric CC[${fixed},${moving},1,4] \ 
                 --convergence 100x70x50x20 \
                 --smoothing-sigmas 2x1.0x0.5x0.0mm \
                 --shrink-factors 6x4x2x1

@@

Additionally, we refined the labelings for each subject of each cohort
using the multi-atlas label fusion algorithm (MALF) developed by \cite{wang2012}
which is also distributed with the ANTs repository.
For a given subject within a data set, every other subject was mapped to 
that subject using the pseudo-geodesic transform.  The set of transformed labelings 
were then used to determine a consensus
labeling for that subject.
This was to minimize the obvious {\it observer dimensionality bias} where manual
raters observe and label in a single dimension at a time.  This
is most easily seen in the axial or sagittal views of the different cohorts 
as labelings were done primarily in the coronal view (see Figure \ref{fig:labels}).  
We include both sets of results.  This provides two sets of labels per subject
for evaluation.

@@

The differences between the two algorithms consist of the Gaussian and 
B-spline parameters governing the update field smoothing, $S_v$.  In our 
experience, smoothing of the total field did not improve the results, at
least for these data (which conforms with our experience with other
data), so the total field smoothing, $S_{\phi}$ is 0 for
both registration approaches.  Specifically, the chosen parameters
for the SyN algorithm were:  $S_{\phi} = \mathcal{N}(0,0)$ and 
$S_{v} = \mathcal{N}(0,3)$ in voxel terms.%
\footnote{
 Or, in equivalent {\tt antsRegistration} command line parlance, {\tt -t
 SyN[0.1,3,0]}.
 }  Although our experience with B-spline SyN is much more limited, we 
were able to choose comparable parameters based on a knot spacing
for the update field of 26 mm at the base level which is reduced by
a factor of two for each subsequent multiresolution level.  This 
yields a final knot spacing of 3.25 mm.%
\footnote{
 The {\tt antsRegistration} command line currently 
 permits specification of the
 B-spline total and update field smoothing only in terms of total mesh size over
 the entire image domain, i.e. {\tt -t BSplineSyN[0.1,10x10x8,0]}.  However, 
 in order to minimize anisotropic smoothing, we determined for each subject 
 the mesh size closest to producing a knot spacing of 26 mm which was typically
 around 10 mesh elements per dimension.  We are currently planning on modifying
 the command line to permit specification of a knot size (similar to the 
 ANTs-based N4 bias correction algorithm \citep{tustison2010} which also uses the
 same underlying B-spline computational machinery).
 }
As specified in the supplementary material associated with \cite{klein2009}, 
this is similar to the 
gradient smoothing parameter for the IRTK FFD algorithm
which also used four multi-resolution levels with an initial knot spacing of
20 mm per dimension for a final knot spacing of 2.5 mm.

>>>>

antsRegistration --dimensionality 3 \
                 --output ${prefix} \
                 --use-histogram-matching 1 \
                 --initial-moving-transform [${fixed},${moving},1] \
                 --transform Rigid[0.1] \
                 --metric MI[${fixed},${moving},1,32,Regular,0.25] \ 
                 --convergence 1000x500x250 \
                 --smoothing-sigmas 2.0x1.0x0.0mm \
                 --shrink-factors 4x3x2 \
                 --transform Affine[0.1] \
                 --metric MI[${fixed},${moving},1,32,Regular,0.25] \ 
                 --convergence 1000x500x250 \
                 --smoothing-sigmas 2.0x1.0x0.0mm \
                 --shrink-factors 4x3x2 \
                 --transform BSplineSyN[0.1,26,0x0x0,3] \
                 --metric CC[${fixed},${moving},1,4] \ 
                 --convergence 100x70x50x20 \
                 --smoothing-sigmas 2x1.0x0.5x0.0mm \
                 --shrink-factors 6x4x2x1

@@

Additionally, we refined the labelings for each subject of each cohort
using the multi-atlas label fusion algorithm (MALF) developed by \cite{wang2012}
which is also distributed with the ANTs repository.
For a given subject within a data set, every other subject was mapped to 
that subject using the pseudo-geodesic transform.  The set of transformed labelings 
were then used to determine a consensus
labeling for that subject.
This was to minimize the obvious {\it observer dimensionality artifacts} where manual
raters observe and label in a single dimension at a time.  This
is most easily seen in the axial or sagittal views of the different cohorts 
as labelings were done primarily in the coronal view (see Figure \ref{fig:labels}).  
We include both sets of results.  This provides two sets of labels per subject
for evaluation.

@@

The differences between the two algorithms consist of the Gaussian and 
B-spline parameters governing the update field smoothing, $S_v$.  In our 
experience, smoothing of the total field did not improve the results, at
least for these data (which conforms with our experience with other
data), so the total field smoothing, $S_{\phi}$ is 0 for
both registration approaches.  Specifically, the chosen parameters
for the SyN algorithm were:  $S_{\phi} = \mathcal{N}(0,0)$ and 
$S_{v} = \mathcal{N}(0,3)$ in voxel terms.%
\footnote{
 Or, in equivalent {\tt antsRegistration} command line parlance, {\tt -t
 SyN[0.1,3,0]}.
 }  Although our experience with B-spline SyN is much more limited, we 
were able to choose comparable parameters based on a knot spacing
for the update field of 26 mm at the base level which is reduced by
a factor of two for each subsequent multiresolution level.  This 
yields a final knot spacing of 3.25 mm.%
\footnote{
 Or, equivalently, {\tt -t
 BSplineSyN[0.1,26,0]}.
% The {\tt antsRegistration} command line currently 
% permits specification of the
% B-spline total and update field smoothing only in terms of total mesh size over
% the entire image domain, i.e. {\tt -t BSplineSyN[0.1,10x10x8,0]}.  However, 
% in order to minimize anisotropic smoothing, we determined for each subject 
% the mesh size closest to producing a knot spacing of 26 mm which was typically
% around 10 mesh elements per dimension.  We are currently planning on modifying
% the command line to permit specification of a knot size (similar to the 
% ANTs-based N4 bias correction algorithm \citep{tustison2010} which also uses the
% same underlying B-spline computational machinery).
 }
As specified in the supplementary material associated with \cite{klein2009}, 
this is similar to the 
gradient smoothing parameter for the IRTK FFD algorithm
which also used four multi-resolution levels with an initial knot spacing of
20 mm per dimension for a final knot spacing of 2.5 mm.
